---
title: "Review 2015"
output:
  html_document:
    theme: united
  pdf_document: default
  word_document: default
---
`r as.character(format(Sys.Date(), format="%B %d, %Y"))`

------
This is Review of 2015

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(MASS)
library(rms)
library(ggplot2)
library(plyr)
library(scales)

rm(list=ls())    #remove all datasets
# setwd("../Data")    #set path

matched <- read.csv("../Data/DataSetPerformance1215.csv", header = T)
matched <- subset(matched, R_2015 != "")
matched$Education <- as.character(matched$Education)
matched$Education[which(matched$Education == "")] <- "High School or Others" #assume blank to high school level
matched$Education <- as.factor(matched$Education)

mydata <- data.frame(matched,"Edu"=numeric(nrow(matched)),stringsAsFactors=FALSE)
mydata$Edu[which(matched$Education == "PhD Level")] <- 4
mydata$Edu[which(matched$Education == "Master Level")] <- 3
mydata$Edu[which(matched$Education == "Bachelor Level")] <- 2
mydata$Edu[which(matched$Education == "Associate Level")] <- 1
mydata$Edu[which(matched$Education == "High School or Others")] <- 0

# creat variables
Gender <- as.numeric(matched$Gender == "M") #dummy Gender
M_Status <- as.numeric(matched$Marital_Status == "Married")  #dummy Married
Active <- as.numeric(matched$Employee_Status_Type == "Active") #dummy Active
R <- as.numeric(matched$R_2015)
Review <- as.factor(matched$R_2015)
Edu <- as.numeric(mydata$Edu)

var <- c("Payroll_Name","Age","Children","VS","Department")
mydata <- data.frame(Review,matched[var],Gender,M_Status,Active,Edu,R)
attach(mydata)

Y <- cbind(Review)
X <- cbind(Gender,Age,M_Status,Children,VS,Active,Edu)
Xvar <- c("Gender","Age","M_Status","Children","VS","Active","Edu")
table(Review)
summary(cbind(X,R))

ddist <- datadist(Xvar)
options(datadist = 'ddist')
ologit <- lrm(Y ~ X, data=mydata)
print(ologit)

```

<hr>

# Overall Review Plot for 2015
```{r Review plot, echo=FALSE, message=FALSE, warning=FALSE}
# Overall Review Plot for matched only
qplot(Review, data=mydata, geom="bar",
      fill=Review) + labs(x = "Review", y = "")
```

<hr>
<!-- results='asis', tidy=TRUE} -->

# Review Scores by Departments
```{r Departments, echo=FALSE, message=FALSE, warning=FALSE}

# library(googleVis)
# op <- options(gvis.plot.tag='chart')  #plot option in Rmd
# 
# dc <- ddply(mydata,.(Department),summarise,median=median(R),mean=mean(R),count=sum(!is.na(Review)),percent=percent(sum(!is.na(Review))/nrow(matched)))
# 
# table <- gvisTable(dc[,1:4],
#                    formats = list(median="#", mean="#.##"),
#                    options = list(width='100%'))
# plot(table)

ddply(mydata,.(Department),summarise,median=median(R),mean=mean(R),count=sum(!is.na(Review)),percent=percent(sum(!is.na(Review))/nrow(mydata)))

# department by scores hist
ggplot(mydata, aes(x=Review,fill=matched$Department)) + geom_bar() + labs(y = "") + theme(legend.title=element_blank()) # this line revome the legend title

# department by scores conditional dist
ggplot(mydata, aes(x=Review,fill=matched$Department)) + geom_bar(position = "fill") + labs(y = "") + theme(legend.title=element_blank()) # this line revome the legend title

```

<hr>

# Review vs Age
No significant pattern

```{r Age, echo=FALSE, message=FALSE, warning=FALSE}
# Age: Hist and mean label
ggplot(mydata, aes(x=Age, fill=Review)) + geom_histogram(binwidth = 5) + labs(y = "") +
  annotate("text", label = paste ("Average Age =",round(mean(mydata$Age),1)),
           x = 55, y = 40, family="serif", fontface="italic", colour="darkred", size=6) +
  annotate("text", label = paste ("Median Age =",round(median(mydata$Age),1)),
           x = 55, y = 37, family="serif", fontface="italic", colour="darkred", size=6)
# Review vs Age
ggplot(mydata, aes(Age, ..count..)) + labs(y = "") +
  geom_density(position = "fill", aes(fill = Review, colour = Review))
```

<hr>

# Review vs Children
No significant pattern

```{r Children, echo=FALSE, message=FALSE, warning=FALSE}
# Children: Hist and mean label
ggplot(mydata, aes(x=Children, fill=Review)) + geom_histogram(binwidth = 1) + labs(y = "") +
  annotate("text", label = paste ("Average Children =",round(mean(mydata$Children),1)),
           x = 3.5, y = 125, family="serif", fontface="italic", colour="darkred", size=6)
# Review vs Children
ggplot(mydata, aes(x=factor(Children), fill=Review)) +
  geom_bar(position="fill") + labs(x = "Children",y = "")
```

<hr>

# Higher Education, better review scores
```{r Education, echo=FALSE, message=FALSE, warning=FALSE}
# education: Hist and Mean label
ggplot(mydata, aes(x=matched$Education,fill=Review)) + geom_bar() + labs(x = "Education",y = "") + scale_x_discrete(limits=c("High School or Others","Associate Level","Bachelor Level","Master Level","PhD Level"))

# Review vs all education
ggplot(mydata, aes(x=matched$Education, fill=Review)) +
  geom_bar(position="fill") + labs(x = "Education",y = "") +
  scale_x_discrete(limits=c("High School or Others","Associate Level","Bachelor Level","Master Level","PhD Level"))

# difference in means between education
ddply(mydata,.(Edu),summarise,median=median(R),mean=mean(R),count=sum(!is.na(Review)),percent=percent(sum(!is.na(Review))/nrow(mydata)))

# department by education hist
ggplot(mydata, aes(x=matched$Education,fill=matched$Department)) + geom_bar() + labs(y = "",x = "Education") + scale_x_discrete(limits=c("High School or Others","Associate Level","Bachelor Level","Master Level","PhD Level")) + theme(legend.title=element_blank()) # this line revome the legend title

# department by education conditional dist
ggplot(mydata, aes(x=matched$Education,fill=matched$Department)) + geom_bar(position = "fill") + labs(y = "",x = "Education") + scale_x_discrete(limits=c("High School or Others","Associate Level","Bachelor Level","Master Level","PhD Level")) + theme(legend.title=element_blank()) # this line revome the legend title
```

<hr>

# Gender F(0) has higher Review
```{r Gender, echo=FALSE, message=FALSE, warning=FALSE}
# Gender: Hist and Mean label
ggplot(mydata, aes(matched$Gender, fill=Review)) + geom_bar() + labs(x = "Gender", y = "") +
  annotate("text", label = paste ("Female:",percent((1-mean(Gender)))),
           x = 1, y = 75, family="serif", fontface="italic", colour="darkred", size=6)
# Gender F has higher Review
ggplot(mydata, aes(x=Review, fill=matched$Gender)) +
  geom_bar(position="fill") + labs(y = "") +
  theme(legend.title=element_blank()) # this line revome the legend title
# difference in means between Gender
ddply(mydata,.(Gender),summarise,median=median(R),mean=mean(R),count=sum(!is.na(Review)))
wilcox.test(R~Gender, alternative="greater") #Nonparametric one side test
ggplot(mydata, aes(x=matched$Gender, fill=Review)) +
  geom_bar(position="fill") + labs(x = "Gender", y = "")

# department by Gender hist
ggplot(mydata, aes(x=matched$Gender,fill=matched$Department)) + geom_bar() + labs(y = "",x = "Gender") + theme(legend.title=element_blank()) # this line revome the legend title

# department by Gender conditional dist
ggplot(mydata, aes(x=matched$Gender,fill=matched$Department)) + geom_bar(position = "fill") + labs(y = "",x = "Gender") + theme(legend.title=element_blank()) # this line revome the legend title

```

<hr>

# Married employees have higher Review
```{r Married, echo=FALSE, message=FALSE, warning=FALSE}
# marital status: Hist and Mean label
ggplot(mydata, aes(x=matched$Marital_Status,fill=Review)) + geom_bar() + labs(x = "Marital Status",y = "") + annotate("text", label = paste ("Marriage Rate:",percent(mean(mydata$M_Status))), x = 3, y = 160, family="serif", fontface="italic", colour="darkred", size=6)
# Review vs all marital status
ggplot(mydata, aes(x=matched$Marital_Status, fill=Review)) +
  geom_bar(position="fill") + labs(x = "Marital Status",y = "")

# difference in means between Married or not
ddply(mydata,.(M_Status),summarise,median=median(R),mean=mean(R),count=sum(!is.na(Review)))
wilcox.test(R~M_Status, alternative="less") #Nonparametric one side test
ggplot(mydata, aes(x=factor(M_Status), fill=Review)) +
  geom_bar(position="fill") + labs(x = "Marital Status: 1 = Married",y = "")

# # department by married
# ggplot(mydata, aes(x=factor(M_Status),fill=matched$Department)) + geom_bar() + labs(y = "",x = "Gender") + theme(legend.title=element_blank()) # this line revome the legend title
# 
# # department by married dist
# ggplot(mydata, aes(x=factor(M_Status),fill=matched$Department)) + geom_bar(position = "fill") + labs(y = "",x = "Gender") + theme(legend.title=element_blank()) # this line revome the legend title
```

<hr>

# Review vs VS
Longer VS, higher Review

```{r VS, echo=FALSE, message=FALSE, warning=FALSE}
# VS: Hist and Mean, median label
ggplot(mydata, aes(x=VS,fill=Review)) + geom_histogram() + labs(y = "") +
  annotate("text", label = paste ("Average VS =",round(mean(mydata$VS),1)),
           x = 15, y = 40, family="serif", fontface="italic", colour="darkred", size=6) +
  annotate("text", label = paste ("Median VS =",round(median(mydata$VS),1)),
           x = 15, y = 37, family="serif", fontface="italic", colour="darkred", size=6)
# Review vs VS
# mydatavs <- subset(mydata, VS != "32.17") #to keep consistent: remove the only VS 30 - Review 2 point
ggplot(mydata, aes(VS, ..count..)) + labs(y = "") +
  geom_density(position = "fill", aes(fill = Review, colour = Review))
```

<hr>

# Actives have higher Review
```{r Active, echo=FALSE, message=FALSE, warning=FALSE}
# Active: Hist and Mean label
ggplot(mydata, aes(matched$Employee_Status_Type, fill=factor(Review))) + geom_bar() + labs(x = "", y = "") +
  annotate("text", label = paste ("Active:",percent((mean(Active)))),
           x = 1, y = 260, family="serif", fontface="italic", colour="darkred", size=6) +
  scale_fill_discrete(guide = guide_legend(title = "Review")) # title text
# difference in means between Active or not (No retired)
ddply(matched,.(Employee_Status_Type),summarise,median=median(R_2015),mean=mean(R_2015),count=sum(!is.na(R_2015)))
wilcox.test(matched$R_2015~matched$Employee_Status_Type, alternative="greater")
ggplot(mydata, aes(x=matched$R_2015, fill=matched$Employee_Status_Type)) +
  geom_bar(position="fill") + labs(y = "", x = "Review") +
  theme(legend.title=element_blank()) # this line revome the legend title
```

<hr>
