---
title: "Review 2015 - Research Department"
output:
  html_document:
    theme: united
  pdf_document: default
  word_document: default
---
`r as.character(format(Sys.Date(), format="%B %d, %Y"))`

------
There are 38 valid review reports for year 2015, average score is 3.184 (higher than 2015 overall review score 2.902). Most of reviews have score 3 - Commendable, and research department donot have review scores under 2 - Satisfactory.

May because the sample size is small, using the same regression model as other review analysis does not generate a good result.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(MASS)
library(rms)
library(ggplot2)
library(plyr)
library(scales)

rm(list=ls())    #remove all datasets
# setwd("../Data")    #set path

matched <- read.csv("../Data/PerformanceRD.csv", header = T)
matched <- subset(matched, R_2015 != "")
matched$Education <- as.character(matched$Education)
matched$Education[which(matched$Education == "")] <- "High School or Others" #assume blank to high school level
matched$Education <- as.factor(matched$Education)

mydata <- data.frame(matched,"Edu"=numeric(nrow(matched)),stringsAsFactors=FALSE)
mydata$Edu[which(matched$Education == "PhD Level")] <- 4
mydata$Edu[which(matched$Education == "Master Level")] <- 3
mydata$Edu[which(matched$Education == "Bachelor Level")] <- 2
mydata$Edu[which(matched$Education == "Associate Level")] <- 1
mydata$Edu[which(matched$Education == "High School or Others")] <- 0

# creat variables
Gender <- as.numeric(matched$Gender == "M") #dummy Gender
M_Status <- as.numeric(matched$Marital_Status == "Married")  #dummy Married
Active <- as.numeric(matched$Employee_Status_Type == "Active") #dummy Active
R <- as.numeric(matched$R_2015)
Review <- as.factor(matched$R_2015)
Edu <- as.numeric(mydata$Edu)
MBO <- as.numeric(matched$MBO)

var <- c("Payroll_Name","Age","Children","VS","Sub_Department")
mydata <- data.frame(Review,matched[var],Gender,M_Status,Active,Edu,MBO,R)
attach(mydata)

Y <- cbind(Review)
X <- cbind(Gender,Age,M_Status,Children,VS,Edu,MBO)
Xvar <- c("Gender","Age","M_Status","Children","VS","Edu","MBO")
table(Review)
summary(cbind(X,R))

ddist <- datadist(Xvar)
options(datadist = 'ddist')
ologit <- lrm(Y ~ X, data=mydata)
print(ologit)

summary(lm(R ~ X))

```

<hr>

# Overall Review Plot for 2015
```{r Review plot, echo=FALSE, message=FALSE, warning=FALSE}
# Overall Review Plot for matched only
qplot(Review, data=mydata, geom="bar", fill=Review) + labs(x = "Review", y = "")
```

<hr>

# Review Scores by Sub_Departments

```{r Sub_Departments, echo=FALSE, message=FALSE, warning=FALSE}

dc <- ddply(mydata,.(Sub_Department),summarise,median=median(R),mean=mean(R),count=sum(!is.na(Review)),percent=percent(sum(!is.na(Review))/nrow(mydata)))
dc <- dc[order(dc$mean, decreasing = T),]
rownames(dc) <- NULL
dc

# boxplot
ggplot(mydata, aes(x=matched$Sub_Department,R)) + geom_boxplot() + coord_flip() + labs(x= "", y = "Review")

# Sub_Department by scores hist
ggplot(mydata, aes(x=Review,fill=matched$Sub_Department)) + geom_bar() + labs(y = "") + theme(legend.title=element_blank()) # this line revome the legend title

# Sub_Department by scores conditional dist
ggplot(mydata, aes(x=Review,fill=matched$Sub_Department)) + geom_bar(position = "fill") + labs(y = "") + theme(legend.title=element_blank()) # this line revome the legend title

```

```{r, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}

r results='asis', eval=FALSE, message=FALSE, warning=FALSE, include=FALSE, tidy=TRUE echo=FALSE}
library(googleVis)
op <- options(gvis.plot.tag='chart')  #plot option in Rmd

dc <- ddply(mydata,.(Sub_Department),summarise,median=median(R),mean=mean(R),count=sum(!is.na(Review)),percent=percent(sum(!is.na(Review))/nrow(matched)))

table <- gvisTable(dc[,1:4],
                   formats = list(median="#", mean="#.##"),
                   options = list(width='100%'))
plot(table)

```

<hr>

# Review vs Age

```{r Age, echo=FALSE, message=FALSE, warning=FALSE}
# Age: Hist and mean label
ggplot(mydata, aes(x=Age, fill=Review)) + geom_histogram(binwidth = 5) + labs(y = "") + annotate("text", label = paste ("Average Age =",round(mean(mydata$Age),1)), x = 55, y = 10, family="serif", fontface="italic", colour="darkred", size=6) + annotate("text", label = paste ("Median Age =",round(median(mydata$Age),1)), x = 55, y = 9, family="serif", fontface="italic", colour="darkred", size=6)

subset(mydata, R == 2)
# # Review vs Age
# ggplot(mydata, aes(Age, ..count..)) + labs(y = "") +
#   geom_density(position = "fill", aes(fill = Review, colour = Review))
```

<hr>

# Review vs Children

```{r Children, echo=FALSE, message=FALSE, warning=FALSE}
# Children: Hist and mean label
ggplot(mydata, aes(x=Children, fill=Review)) + geom_histogram(binwidth = 1) + labs(y = "") + annotate("text", label = paste ("Average Children =",round(mean(mydata$Children),1)), x = 3, y = 15, family="serif", fontface="italic", colour="darkred", size=6)
# # Review vs Children
# ggplot(mydata, aes(x=factor(Children), fill=Review)) +
#   geom_bar(position="fill") + labs(x = "Children",y = "")
```

<hr>

# Review vs Education

The analysis has assigned 5 levels of education: 0 – High School or Others, 1 – Associate Level, 2 – Bachelor Level, 3 – Master Level, 4 – PhD Level. 

```{r Education, echo=FALSE, message=FALSE, warning=FALSE}
# difference in means between education
ddply(mydata,.(Edu),summarise,median=median(R),mean=mean(R),count=sum(!is.na(Review)),percent=percent(sum(!is.na(Review))/nrow(mydata)))

ggplot(mydata, aes(x=matched$Education,R)) + geom_boxplot() + labs(x= "", y = "Review") + scale_x_discrete(limits=c("High School or Others","Associate Level","Bachelor Level","Master Level","PhD Level"))

# education: Hist and Mean label
ggplot(mydata, aes(x=matched$Education,fill=Review)) + geom_bar() + labs(x = "Education",y = "") + scale_x_discrete(limits=c("High School or Others","Associate Level","Bachelor Level","Master Level","PhD Level"))

# # Review vs all education
# ggplot(mydata, aes(x=matched$Education, fill=Review)) +
#   geom_bar(position="fill") + labs(x = "Education",y = "") +
#   scale_x_discrete(limits=c("High School or Others","Associate Level","Bachelor Level","Master Level","PhD Level"))

# Sub_Department by education hist
ggplot(mydata, aes(x=matched$Education,fill=matched$Sub_Department)) + geom_bar() + labs(y = "",x = "Education") + scale_x_discrete(limits=c("High School or Others","Associate Level","Bachelor Level","Master Level","PhD Level")) + theme(legend.title=element_blank()) # this line revome the legend title

# # Sub_Department by education conditional dist
# ggplot(mydata, aes(x=matched$Education,fill=matched$Sub_Department)) + geom_bar(position = "fill") + labs(y = "",x = "Education") + scale_x_discrete(limits=c("High School or Others","Associate Level","Bachelor Level","Master Level","PhD Level")) + theme(legend.title=element_blank()) # this line revome the legend title
```

<hr>

# Review vs Gender

```{r Gender, echo=FALSE, message=FALSE, warning=FALSE}
# Gender: Hist and Mean label
ggplot(mydata, aes(matched$Gender, fill=Review)) + geom_bar() + labs(x = "Gender", y = "") + annotate("text", label = paste ("Female:",percent((1-mean(Gender)))), x = 1, y = 23, family="serif", fontface="italic", colour="darkred", size=6)
# Gender F has higher Review
ggplot(mydata, aes(x=Review, fill=matched$Gender)) + geom_bar(position="fill") + labs(y = "") + theme(legend.title=element_blank()) # this line revome the legend title
# difference in means between Gender
ddply(mydata,.(Gender),summarise,median=median(R),mean=mean(R),count=sum(!is.na(Review)))
wilcox.test(R~Gender, alternative="greater") #Nonparametric one side test
ggplot(mydata, aes(x=matched$Gender, fill=Review)) + geom_bar(position="fill") + labs(x = "Gender", y = "")

# Sub_Department by Gender hist
ggplot(mydata, aes(x=matched$Gender,fill=matched$Sub_Department)) + geom_bar() + labs(y = "",x = "Gender") + theme(legend.title=element_blank()) # this line revome the legend title

# Sub_Department by Gender conditional dist
ggplot(mydata, aes(x=matched$Gender,fill=matched$Sub_Department)) + geom_bar(position = "fill") + labs(y = "",x = "Gender") + theme(legend.title=element_blank()) # this line revome the legend title

```

<hr>

# Review vs Marital_Status

```{r Married, echo=FALSE, message=FALSE, warning=FALSE}
# marital status: Hist and Mean label
ggplot(mydata, aes(x=matched$Marital_Status,fill=Review)) + geom_bar() + labs(x = "Marital Status",y = "") + annotate("text", label = paste ("Marriage Rate:",percent(mean(mydata$M_Status))), x = 2.5, y = 25, family="serif", fontface="italic", colour="darkred", size=6)
# # Review vs all marital status
# ggplot(mydata, aes(x=matched$Marital_Status, fill=Review)) +
#   geom_bar(position="fill") + labs(x = "Marital Status",y = "")

# difference in means between Married or not
ddply(mydata,.(M_Status),summarise,median=median(R),mean=mean(R),count=sum(!is.na(Review)))
wilcox.test(R~M_Status, alternative="less") #Nonparametric one side test
ggplot(mydata, aes(x=factor(M_Status), fill=Review)) + geom_bar(position="fill") + labs(x = "Marital Status: 1 = Married",y = "")

# # Sub_Department by married
# ggplot(mydata, aes(x=factor(M_Status),fill=matched$Sub_Department)) + geom_bar() + labs(y = "",x = "Gender") + theme(legend.title=element_blank()) # this line revome the legend title
# 
# # Sub_Department by married dist
# ggplot(mydata, aes(x=factor(M_Status),fill=matched$Sub_Department)) + geom_bar(position = "fill") + labs(y = "",x = "Gender") + theme(legend.title=element_blank()) # this line revome the legend title
```

<hr>

# Review vs Vesting Service

```{r VS, echo=FALSE, message=FALSE, warning=FALSE}
# VS: Hist and Mean, median label
ggplot(mydata, aes(x=VS,fill=Review)) + geom_histogram() + labs(y = "") + annotate("text", label = paste ("Average VS =",round(mean(mydata$VS),1)), x = 15, y = 7, family="serif", fontface="italic", colour="darkred", size=6) + annotate("text", label = paste ("Median VS =",round(median(mydata$VS),1)), x = 15, y = 6.5, family="serif", fontface="italic", colour="darkred", size=6)
# Review vs VS
# mydatavs <- subset(mydata, VS != "32.17") #to keep consistent: remove the only VS 30 - Review 2 point
subset(mydata, R == 2)

```

<hr>
